---
title: "parallel_computaion"
author: "Jean Remy"
date: "January 17, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install and load other packages using pacman

```{r, }
library(pacman) 
pacman::p_load(furrr, purrr, tidyverse, tictoc)
```

## library(purr)

Computations done sequentially. If the next operation dedends on the previous ones.

### library(furr)

Computations are done in parallel (at the same time). The tasks are independent

# The plan() function 

This is how you set the processing type of the future_(): 

*  Default is sequential : furrr works exactly like purrr
*  All we need to do is change our plan() depending on our compute architecture : tell the computer to execute tasks 
   in parallel
*  On your local computer, use plan(multiprocess)

## initiate the plan

```{r, }
no_cores <- availableCores() - 1# leave one core for the other software on your computer to use.
no_cores
plan(multiprocess, workers = no_cores) 
```
 

# SET UP YOUR CORES

With library(furrr) we don’t have to worry about sending our special functions, packages, 
and data to each process before start our multiprocess code. 
This is one of my favorite features! No longer do we have to

# LOAD LIBRARIES

```{r, echo = FALSE }
#devtools::install_github("hathawayj/buildings")

library(buildings) # remember that the 'permits' data object is created when the library is loaded.
a <- 4
ff <- function(x){
  for (i in 1:1000){
    i
  }
  
  ggplot() + geom_point(x = permits[x, "value"])
}
list_object <- as.list(1:7500)
list_object

```


# UNDERSTANDING map() and future_map()

Now the magic works best when we think with list objects. 
List objects are a great way to use multiple processes in a divide/recombine fashion.

```{r, }
tic()
temp1 <- map(list_object, ff)
toc()

tic()
temp1 <- future_map(list_object, ff)
toc()
```

As library(future) only starts processes when needed we don’t have to worry about stopping the cluster.

# Additional Example

The future_map() functions work the same as the map() functions if you set plan(sequential). See the example below.

```{r, }

second_sequence <- rep(5, 8)

tic()
nothingness <- map(second_sequence, ~Sys.sleep(.x))
toc()

plan(sequential)
tic()
nothingness <- future_map(second_sequence, ~Sys.sleep(.x))
toc()

plan(multiprocess)
tic()
nothingness <- future_map(second_sequence, ~Sys.sleep(.x))
toc()
```


[](https://byuistats.github.io/M335/parallel_furrr.html)


